Constructing a universe for the setoid model

### 2020-07-04 paper.tex
Small corrections

Type theory with equality reflection, also known as \emph{extensional type
  theory} (ETT) does support extensional reasoning to some degree, but
unfortunately equality reflection makes the problem of type-checking ETT terms
computationally unfeasible, if not impossible.

On the other hand, \emph{intensional type theory} (ITT) has nice computational
properties like decidable type checking that can make it more suitable for
computer implementation, but as normally defined (for example, in \cite{mltt75})
it severely lacks extensionality.
%
It is known from model constructions that extensional principles like funext are
consistent with ITT.
%
Moreover, ITT extended with the principle of \emph{uniqueness of identity
  proofs} (UIP) and funext is known to be as powerful as ETT
\cite{conservativity}. We could recover the expressive power of ETT by adding
these principles to ITT as axioms, however destroying some computational
properties like canonicity.

@@

A first fundamental consideration has to with the very definition of the setoid
universe: as any type in the setoid model, this universe must be a setoid and
thus come equipped with an equivalence relation. However, unlike the universe of
propositions, a universe of setoids cannot be univalent, since this would force
it to be a groupoid. The obvious choice is therefore to have a non-univalent
universe, and instead define the universe's relation so that it reflects a
simple syntactic equality of codes rather than setoid equivalence.

@@

In the definition above, $p$ has type $\|| \Sigma (R: \El\,A \to \El\,B \to
\mProp)(...) \||$. If the type wasn't propositionally truncated, we could define
$p \vdash a \sim_{\El} b$ by extracting the relation out of the first component
of $p$, and apply it to $a, b$. That is, $p \vdash a \sim_{\El} b :\equiv
\pi_1\,p\,a\,b$. This would make the definition of $\_\sim_\setoidU\_$ and
$\_\vdash\_\sim_{\El}\_$ in line with how we defined $\setoidU$ and $\El$.

>>>>

Type theory with equality reflection, also known as \emph{extensional type
  theory} (ETT) does support extensional reasoning to some degree, but
unfortunately equality reflection makes the problem of type-checking ETT terms
computationally unfeasible: it is undecidable.

On the other hand, \emph{intensional type theory} (ITT) has nice computational
properties like decidable type checking that can make it more suitable for
computer implementation, but as usually defined (for example, in \cite{mltt75})
it severely lacks extensionality.
%
It is known from model constructions that extensional principles like funext are
consistent with ITT.
%
Moreover, ITT extended with the principle of \emph{uniqueness of identity
  proofs} (UIP) and funext is known to be as powerful as ETT
\cite{conservativity}. We could recover the expressive power of ETT by adding
these principles to ITT as axioms, however destroying some computational
properties like canonicity.

@@

A first fundamental consideration has to do with the very definition of the setoid
universe: as any type in the setoid model, this universe must be a setoid and
thus come equipped with an equivalence relation. However, unlike the universe of
propositions, a universe of setoids cannot be univalent, since this would force
it to be a groupoid. The obvious choice is therefore to have a non-univalent
universe, and instead define the universe's relation so that it reflects a
simple syntactic equality of codes rather than setoid equivalence.

@@

In the definition above, $p$ has type $\parallel \Sigma\ (R: \El\,A \to \El\,B \to
\mProp)\ (...) \parallel$. If the type wasn't propositionally truncated, we could define
$p \vdash a \sim_{\El} b$ by extracting the relation out of the first component
of $p$, and apply it to $a, b$. That is, $p \vdash a \sim_{\El} b :\equiv
\pi_1\,p\,a\,b$. This would make the definition of $\_\sim_\setoidU\_$ and
$\_\vdash\_\sim_{\El}\_$ in line with how we defined $\setoidU$ and $\El$.
